```{r knitrSetup}
# sets output for inline results nicely
knit_hooks$set(inline = identity) 

# kills knitr if there is an error, therefore we don't waste time generating error messages
knit_hooks$set(error = function(x, options) stop(x)) 
```

```{r rsetup}
# these are extremely useful options for working in R
options(save.defaults=list(compress="xz"), stringsAsFactors=FALSE)
```

# MAQC AbsIDConvert

## Abstract

Current transcriptomics analyses are hampered by an inability to easily convert between various types of identifiers,
making it difficult to easily compare the results of studies performed using different types of platforms. AbsIDConvert,
a tool recently developed by the Rouchka group, makes interconversion of identifiers possible by mapping each type of
identifier to a genomic interval, and using overlapping intervals as the basis for the conversion. Using the MicroArray
Quality Control (MAQC) dataset as an example, this work will demonstrate how the agreement between different experiments 
changes depending on how the platform specific identifiers are converted to allow comparison.